---
title: "Astrocyte_paper_figures"
output: html_document
date: "2024-08-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(edgeR)
library(RColorBrewer)
library(viridis)
library(dplyr)
library(tidyr)
```


Figure 5E - per-cluster marker expression to highlight astrocyte subtypes

```{r}
agg_cl <- readRDS("/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/DEGs/aggregated_cells.AD_MS_PD.070724.DonoruniqueID_and_cluster.SE.rds") ## cain donors are fixed here


get_plot_for_gene <- function(agg_cl, genename) {

  gene = rowData(agg_cl)[rowData(agg_cl)$symbol == genename,]$ID
  
  df_mic <- data.frame(expr = t(assay(agg_cl[gene,],2)), cluster = agg_cl$cluster)
                      
  colors <- c(adjustcolor('#31a354', alpha.f = 0.9), '#8856a7', adjustcolor('#fa9fb5', 0.9),  '#2b8cbe',    adjustcolor('#fec44f', alpha.f = 0.9), adjustcolor('#c51b8a', 0.9),  adjustcolor('#d95f0e', alpha.f = 0.9), adjustcolor('turquoise1', alpha.f = 0.9))
 
 p <- ggplot(df_mic, aes(cluster,!!sym(gene), fill=cluster)) + scale_fill_manual(values = colors) +
     geom_violin(trim=TRUE) +
    theme(legend.position = "none") + theme_classic() + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5)) + ylab("logCPM") + ggtitle(genename)

  return(p)
}


```

Replot cellularity figure using percentages

```{r}

coldata <- readRDS("/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/integration/AD_MS_PD_astrocytes.CLEAN_from_non_astrocytes_AND_clusters_high_in_mito.cellbender_counts.20cell_filter.COLDATA.rds")

coldata$diagnosis_harmonized_by_disease <- 
  ifelse(coldata$studybatch %in% c("jakel", "schirmer", "absinta", "bryois") & coldata$diagnosis_harmonized == "Control", "MS_Control",
         
ifelse(coldata$studybatch %in% c("smajic","wang", "kamath") &
coldata$diagnosis_harmonized == "Control", "PD_Control",

ifelse(coldata$studybatch %in% c("cain", "morabito","gerrits", "smith", "SEA-AD", "liddelow") &
coldata$diagnosis_harmonized == "Control", "AD_Control",
coldata$diagnosis_harmonized)))
coldata$diagnosis_harmonized_by_disease <- ifelse(coldata$diagnosis_harmonized == "RRMS", "MS", coldata$diagnosis_harmonized_by_disease)

#### update coldata file

saveRDS(coldata,"/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/integration/AD_MS_PD_astrocytes.CLEAN_from_non_astrocytes_AND_clusters_high_in_mito.cellbender_counts.20cell_filter.COLDATA.correct_diagnosis_harmonized_by_disease.rds")

coldata <- readRDS("/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/integration/AD_MS_PD_astrocytes.CLEAN_from_non_astrocytes_AND_clusters_high_in_mito.cellbender_counts.20cell_filter.COLDATA.correct_diagnosis_harmonized_by_disease.rds")

#### update coldata file to remove the cain samples that are non-clear AD/Control for downstream testing
cain_donors <- as.data.frame(colData(agg_cl)) %>% dplyr::filter(studybatch == "cain") %>% dplyr::select(donor) %>% unique

coldata.cain_fixed <- as.data.frame(coldata) %>% dplyr::filter(! donor %in% cain_donors$donor)

saveRDS(coldata.cain_fixed,"/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/integration/AD_MS_PD_astrocytes.CLEAN_from_non_astrocytes_AND_clusters_high_in_mito.cellbender_counts.20cell_filter.COLDATA.correct_diagnosis_harmonized_by_disease.Cain_donors_subsetted.rds")

coldata <- readRDS("/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/integration/AD_MS_PD_astrocytes.CLEAN_from_non_astrocytes_AND_clusters_high_in_mito.cellbender_counts.20cell_filter.COLDATA.correct_diagnosis_harmonized_by_disease.Cain_donors_subsetted.rds")

abundances <- table(coldata$clusters_res_0.5,coldata$donorIdUnique)
abundances <- unclass(abundances) 

extra.info <- coldata[match(colnames(abundances), coldata$donorIdUnique),]
d <- DGEList(abundances, samples=extra.info)
d = calcNormFactors(d)
d= estimateCommonDisp(d, verbose=TRUE)

norm_counts <- as.data.frame(t(d$counts)) 
colnames(norm_counts) <- paste("Cluster ", colnames(norm_counts), sep="")
norm_counts <- norm_counts %>% mutate(donorIdUnique = rownames(.))

coldata_short <- as.data.frame(coldata) %>% dplyr::select(donorIdUnique,diagnosis_harmonized_by_disease,studybatch,apoe) %>% unique


df_long_final <- left_join(norm_counts,coldata_short, by="donorIdUnique") 
percentages <- data.frame(matrix(nrow = dim(df_long_final)[1],ncol = dim(df_long_final)[2]))
sums <- data.frame(clustersums=rowSums(df_long_final[,1:dim(norm_counts)[2]-1]),ident=df_long_final$donorIdUnique)



for (clust in 1:length(levels(factor(coldata$clusters_res_0.5)))) {
  for (sample in 1:length(df_long_final[,clust])) {
    percent <- (df_long_final[sample,clust]/sums[sample,1])*100
    percentages[sample,clust]<- percent
  }
}


cols2replace <- dim(norm_counts)[2]:(dim(norm_counts)[2]+dim(coldata_short)[2]-1)
percentages[,cols2replace]<- df_long_final[,cols2replace]



colnames(percentages) <-  colnames(df_long_final)
df_long <- percentages %>% 
  pivot_longer(c(c(paste0("Cluster ",1:(dim(norm_counts)[2]-1)))))

level_order <- c("AD_Control","AD","MS_Control","MS","PD_Control","PD")
df_long$diagnosis_harmonized_by_disease <- factor(df_long$diagnosis_harmonized_by_disease,levels = level_order) 

df_long$name <- factor(df_long$name,levels =c(paste0("Cluster ",1:(dim(norm_counts)[2]-1))))

fill.codes <- c("white","darkgreen","white", "purple","white","mediumblue")
color.codes <- c("darkgreen","darkgreen","purple","purple","mediumblue","mediumblue")

zone <- levels(factor(df_long$diagnosis_harmonized_by_disease))

#df_long_no_cl4 <- df_long %>% dplyr::filter()


png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/Human_CellularityPlot_with_percentages.png",
    width = 6000,height=1400,res=300)
ggplot(df_long, aes(x=diagnosis_harmonized_by_disease, y=value,fill=diagnosis_harmonized_by_disease,colour = diagnosis_harmonized_by_disease)) +
  theme_classic() +
  geom_jitter(color="darkgrey",width = 0.2,size=1,alpha=.5)  +
  geom_boxplot(outlier.shape=NA,alpha=0.5)+ 
  
  facet_wrap(~name,scales = "free", ncol=8) + xlab("Disease Label") +
  theme(axis.text.x = element_text(angle=75,vjust = 0.5),axis.text.y = element_text(size=20)) +
  ylab("Cellularity Percent") +  
  theme(strip.text.x = element_text(size = 14,face = "italic"),
        axis.title=element_text(size=14,face="bold"),
        legend.text = element_text(size=14),
        legend.title = element_text(size=14),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  scale_y_continuous(limits = c(0, 100), breaks = c(0, 25, 50,75))+
  scale_fill_manual(values=setNames(fill.codes, zone))+
  scale_color_manual(values=setNames(color.codes, zone))

dev.off()
# 
# 
# # Statistics ---------------------
# 
# sig.clusters.no_clr <- list()
# 
# ## slipping cluster 4 because it's unidentified:
# 
# 
# for(cluster in 1:length(levels(factor(coldata$clusters_res_0.5)))){
#   clusterName <- paste0("Cluster ",levels(factor(coldata$clusters_res_0.5))[cluster])
#   dataCluster <- df_long[which(df_long$name==clusterName),]
#   # res <- kruskal.test(value ~ diagnosis_harmonized_by_disease, data = dataCluster)
#   # # check if there is any significance
#   # if(res$p.value < 0.05/length(levels(factor(coldata$clusters_res_0.5)))){ # bonferroni correction
#   #   sig.clusters[[clusterName]] <- res
#     
#     # do a wilcox test to find which specific groups are different within cluster
#     res.AD <- wilcox.test(dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="AD_Control")],
#                           dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="AD")])
#     res.MS <- wilcox.test(dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="MS_Control")],
#                           dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="MS")])
#     res.PD <- wilcox.test(dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="PD_Control")],
#                           dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="PD")])
#    
#     sig.clusters.no_clr[["AD"]][[clusterName]] <- res.AD
#     sig.clusters.no_clr[["MS"]][[clusterName]] <- res.MS    
#     sig.clusters.no_clr[["PD"]][[clusterName]] <- res.PD
#      
#   }



```

Let's try to CLR adjusted approach
```{r}
abundances <- table(coldata$clusters_res_0.5,coldata$donorIdUnique)
abundances <- unclass(abundances) 

extra.info <- coldata[match(colnames(abundances), coldata$donorIdUnique),]
d <- DGEList(abundances, samples=extra.info)
d = calcNormFactors(d)
d= estimateCommonDisp(d, verbose=TRUE)

norm_counts <- as.data.frame(t(d$counts)) 
colnames(norm_counts) <- paste("Cluster ", colnames(norm_counts), sep="")
norm_counts <- norm_counts %>% mutate(donorIdUnique = rownames(.))

coldata_short <- as.data.frame(coldata) %>% dplyr::select(donorIdUnique,diagnosis_harmonized_by_disease,studybatch,apoe) %>% unique

proportions <- t(apply(norm_counts[,1:8], 1, function(x) (x/sum(x)))) %>% as.data.frame()
value_to_add_to_offset_zeros <- min(proportions[proportions > 0])
proportions <- proportions + value_to_add_to_offset_zeros

CLR_transf_proportions <- t(apply(proportions, 1, function(x) log((x/exp(mean(log(x)))) + 1))) %>% as.data.frame()

CLR_transf_proportions$donorIdUnique <- rownames(CLR_transf_proportions)

CLR_transf_proportions_full <- left_join(CLR_transf_proportions, coldata_short %>% dplyr::select(donorIdUnique, diagnosis_harmonized_by_disease, studybatch, apoe) %>% unique, by = "donorIdUnique")


df_long <- CLR_transf_proportions_full %>% 
  pivot_longer(c(c(paste0("Cluster ",1:(dim(norm_counts)[2]-1)))))

level_order <- c("AD_Control","AD","MS_Control","MS","PD_Control","PD")
df_long$diagnosis_harmonized_by_disease <- factor(df_long$diagnosis_harmonized_by_disease,levels = level_order) 

df_long$name <- factor(df_long$name,levels =c(paste0("Cluster ",1:(dim(norm_counts)[2]-1))))

fill.codes <- c("white","darkgreen","white", "purple","white","mediumblue")
color.codes <- c("darkgreen","darkgreen","purple","purple","mediumblue","mediumblue")

zone <- levels(factor(df_long$diagnosis_harmonized_by_disease))

#df_long_no_cl4 <- df_long %>% dplyr::filter()

png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/Human_CellularityPlot_with_CLR_transformed_proportions.png",
    width = 6000,height=1400,res=300)
ggplot(df_long, aes(x=diagnosis_harmonized_by_disease, y=value,fill=diagnosis_harmonized_by_disease,colour = diagnosis_harmonized_by_disease)) +
  theme_classic() +
  geom_jitter(color="darkgrey",width = 0.2,size=1,alpha=.5)  +
  geom_boxplot(outlier.shape=NA,alpha=0.5)  +
  facet_wrap(~factor(name),scales = "free", ncol=8) + xlab("Disease Label") +
  theme(axis.text.x = element_text(angle=75,vjust = 0.5),axis.text.y = element_text(size=20)) +
  ylab("Cellularity Percent") +  
  theme(strip.text.x = element_text(size = 14,face = "italic"),
        axis.title=element_text(size=14,face="bold"),
        legend.text = element_text(size=14),
        legend.title = element_text(size=14),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  scale_y_continuous(limits = c(0, 7))+
  scale_fill_manual(values=setNames(fill.codes, zone))+
  scale_color_manual(values=setNames(color.codes, zone))
dev.off()


# Statistics ---------------------

sig.clusters <- list()

## slipping cluster 4 because it's unidentified:


for(cluster in 1:length(levels(factor(coldata$clusters_res_0.5)))){
  clusterName <- paste0("Cluster ",levels(factor(coldata$clusters_res_0.5))[cluster])
  dataCluster <- df_long[which(df_long$name==clusterName),]
  res <- kruskal.test(value ~ diagnosis_harmonized_by_disease, data = dataCluster)
  # check if there is any significance
  if(res$p.value < 0.05/length(levels(factor(coldata$clusters_res_0.5)))){ # bonferroni correction
    sig.clusters[[clusterName]] <- res
    
    # do a wilcox test to find which specific groups are different within cluster
    res.AD <- wilcox.test(dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="AD_Control")],
                          dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="AD")])
    res.MS <- wilcox.test(dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="MS_Control")],
                          dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="MS")])
    res.PD <- wilcox.test(dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="PD_Control")],
                          dataCluster$value[which(dataCluster$diagnosis_harmonized_by_disease=="PD")])
   
    sig.clusters[["AD"]][[clusterName]] <- res.AD
    sig.clusters[["MS"]][[clusterName]] <- res.MS    
    sig.clusters[["PD"]][[clusterName]] <- res.PD
     
  }
  
}

saveRDS(sig.clusters, "/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/integration/Sig.clusters_with_CLR_test.rds")
```

With CLR, the following are significant:

1) AD - clusters 3, 5 and 8
2) MS - clusters 1, 5, 6 and 8

Let's see how these look across diseases 

```{r}
get_disease_and_cluster_specific_plot <- function(df_long, disease, cluster, col1) {
  
  df_long.filt <- df_long %>% dplyr::filter(diagnosis_harmonized_by_disease %in% c(paste(disease, "_Control", sep=""), disease)) %>% dplyr::filter(name == cluster)
  
p <- ggplot(df_long.filt, aes(x=diagnosis_harmonized_by_disease, y=value,fill=diagnosis_harmonized_by_disease,colour = diagnosis_harmonized_by_disease)) +
  theme_classic() +
  geom_jitter(color="darkgrey",width = 0.2,size=1,alpha=.5)  +
  geom_boxplot(outlier.shape=NA,alpha=0.5) + xlab("Studies") +
   facet_wrap(~studybatch, scales = "free", ncol=8) + 
  theme(axis.text.x = element_text(angle=75,vjust = 0.5),axis.text.y = element_text(size=20)) +
  ylab("Cellularity Percent") +  
  theme(strip.text.x = element_text(size = 14,face = "italic"),
        axis.title=element_text(size=10,face="italic"),
        legend.text = element_text(size=14),
        legend.title = element_text(size=14),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  scale_color_manual(values = c(col1, col1)) +
  scale_fill_manual(values = c("white", col1)) + ggtitle(paste(disease, cluster, sep=" ")) +   theme(plot.title = element_text(hjust = 0.5))
  
  return(p)
}

cl3_ad <- get_disease_and_cluster_specific_plot(df_long, "AD", "Cluster 3", "darkgreen")
cl5_ad <- get_disease_and_cluster_specific_plot(df_long, "AD", "Cluster 5", "darkgreen")
cl8_ad <- get_disease_and_cluster_specific_plot(df_long, "AD", "Cluster 8", "darkgreen")


png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/AD_sig_clusters_per_study.png",
    width = 4000,height=2000,res=300)
gridExtra::grid.arrange(cl3_ad, cl5_ad, cl8_ad, ncol = 1)
dev.off()

cl1_ms <- get_disease_and_cluster_specific_plot(df_long, "MS", "Cluster 1", "purple")
cl5_ms <- get_disease_and_cluster_specific_plot(df_long, "MS", "Cluster 5", "purple")
cl6_ms <- get_disease_and_cluster_specific_plot(df_long, "MS", "Cluster 6", "purple")
cl8_ms <- get_disease_and_cluster_specific_plot(df_long, "MS", "Cluster 8", "purple")

png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/MS_sig_clusters_per_study.png",
    width = 3000,height=2400,res=300)
gridExtra::grid.arrange(cl1_ms, cl5_ms, cl6_ms, cl8_ms, ncol = 1)
dev.off()



```

Examine cluster markers

```{r}

feat <- as.data.frame(genomitory::getFeatures("GMTY17:GRCh38/GRCh38.IGIS4.0.genes.rds@REVISION-3")) %>% dplyr::select(ID, symbol)

get_cluster_marker_symbols <- function(res, feat) {
  
  m.out <- readRDS(paste("/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/integration/m.out.harmony_2_donor_groupvar.", res,".rds", sep=""))
  m.out <- m.out$statistics
  
  m.out.with_symbols <- list()
  
  for (i in 1:length(m.out)){
  tmp <- m.out[[i]]
  tmp$ID <- rownames(tmp)
  m.out.with_symbols[[i]] <- left_join(tmp, feat, by = "ID")
  }

return(m.out.with_symbols)  
}

m.out.0.5.with_symbols <- get_cluster_marker_symbols(0.5, feat)
names(m.out.0.5.with_symbols) <- paste0("Cluster ",1:8)

```


```{r}

res_ad <- readRDS("/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/DEGs/res_dsl.AD.cain_fixed.072924.rds")

res_ms <- readRDS("/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/DEGs/res_dsl.MS.080124.rds")

res_pd <- readRDS("/gstore/data/omni/neuroscience/AD/Astrocytes_AD_MS_PD_meta/AD_MS_PD_cellbender_integration/harmony_integration/another_run_after_removing_high_mito_clusters/DEGs/res_dsl.PD.072924.rds")



```

Let's run my own pathways and see how they look per disease

```{r}
library(clusterProfiler)
library(enrichplot)

sorted_ad_candidates <- res_ad[order(res_ad$dl_mu,decreasing = TRUE),]
sorted_ad_candidates <- sorted_ad_candidates %>% dplyr::select(dl_mu, symbol) %>% drop_na()
sorted_ad_candidates.for_gsea <- sorted_ad_candidates$dl_mu
names(sorted_ad_candidates.for_gsea) <- sorted_ad_candidates$symbol


sorted_pd_candidates <- res_pd[order(res_pd$dl_mu,decreasing = TRUE),]
sorted_pd_candidates <- sorted_pd_candidates %>% dplyr::select(dl_mu, symbol) %>% drop_na()
sorted_pd_candidates.for_gsea <- sorted_pd_candidates$dl_mu
names(sorted_pd_candidates.for_gsea) <- sorted_pd_candidates$symbol

sorted_ms_candidates <- res_ms[order(res_ms$dl_mu,decreasing = TRUE),]
sorted_ms_candidates <- sorted_ms_candidates %>% dplyr::select(dl_mu, symbol) %>% drop_na()
sorted_ms_candidates.for_gsea <- sorted_ms_candidates$dl_mu
names(sorted_ms_candidates.for_gsea) <- sorted_ms_candidates$symbol


ad_gsea <- gseGO(gene = sorted_ad_candidates.for_gsea,
                      OrgDb = "org.Hs.eg.db",
                      keyType = 'SYMBOL',
                      ont = "ALL",
                      pvalueCutoff = 0.05,
                      minGSSize    = 50,
                      maxGSSize    = 500)
pd_gsea <- gseGO(gene = sorted_pd_candidates.for_gsea,
                      OrgDb = "org.Hs.eg.db",
                      keyType = 'SYMBOL',
                      ont = "ALL",
                      pvalueCutoff = 0.05,
                      minGSSize    = 50,
                      maxGSSize    = 500)
ms_gsea <- gseGO(gene = sorted_ms_candidates.for_gsea,
                      OrgDb = "org.Hs.eg.db",
                      keyType = 'SYMBOL',
                      ont = "ALL",
                      pvalueCutoff = 0.05,
                      minGSSize    = 50,
                      maxGSSize    = 500)


png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/Pathways_combined_AD_only.upregulated.png", res = 300, width = 2600, height = 2200)
dotplot(filter(ad_gsea, NES>1), showCategory = 20) + ggtitle("AD GSEA Pathways Upregulated") + 
  theme(axis.title.x = element_text(size = 14,face="bold"),
        axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(hjust = 0.5,size=16,face=20)) + xlab("")
dev.off()


png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/Pathways_combined_MS_only.upregulated.png", res = 300, width = 2600, height = 2800)
dotplot(filter(ms_gsea, NES>1), showCategory = 20) + ggtitle("MS GSEA Pathways Upregulated") + 
  theme(axis.title.x = element_text(size = 14,face="bold"),
        axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(hjust = 0.5,size=16,face=20)) + xlab("")
dev.off()


png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/Pathways_combined_PD_only.upregulated.png", res = 300, width = 2600, height = 2200)
dotplot(filter(pd_gsea, NES>1), showCategory = 20) + ggtitle("PD GSEA Pathways Upregulated") + 
  theme(axis.title.x = element_text(size = 14,face="bold"),
        axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(hjust = 0.5,size=16,face=20)) + xlab("")
dev.off()

## Grouped Pathways =================================
ck_cc <- compareCluster(list(AD=sorted_ad_candidates.for_gsea,MS=sorted_ms_candidates.for_gsea,PD=sorted_pd_candidates.for_gsea),
                     fun = "gseGO",OrgDb = org.Hs.eg.db::org.Hs.eg.db ,ont="CC",
                      keyType = 'SYMBOL', eps=0)


ck_mf <- compareCluster(list(AD=sorted_ad_candidates.for_gsea,MS=sorted_ms_candidates.for_gsea,PD=sorted_pd_candidates.for_gsea),
                     fun = "gseGO",OrgDb = org.Hs.eg.db::org.Hs.eg.db ,ont="MF",
                      keyType = 'SYMBOL', eps=0)

ck_bp <- compareCluster(list(AD=sorted_ad_candidates.for_gsea,MS=sorted_ms_candidates.for_gsea,PD=sorted_pd_candidates.for_gsea),
                     fun = "gseGO",OrgDb = org.Hs.eg.db::org.Hs.eg.db ,ont="BP",
                      keyType = 'SYMBOL', eps=0)


cc_up <- dotplot(filter(ck_cc, NES>1), by="NES") + ggtitle("GO CC Pathways Upregulated") + 
  theme(axis.title.x = element_text(size = 14,face="bold"),
        axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(hjust = 0.5,size=16,face=20)) + xlab("")

cc_down <- dotplot(filter(ck_cc, NES<1), by="NES") + ggtitle("GO CC Pathways Downregulated") + 
  theme(axis.title.x = element_text(size = 14,face="bold"),
        axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(hjust = 0.5,size=16,face=20)) + xlab("")

mf_up <- dotplot(filter(ck_mf, NES>1), by="NES") + ggtitle("GO MF Pathways Upregulated") + 
  theme(axis.title.x = element_text(size = 14,face="bold"),
        axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(hjust = 0.5,size=16,face=20)) + xlab("")

mf_down <- dotplot(filter(ck_mf, NES<1), by="NES") + ggtitle("GO MF Pathways Downregulated") + 
  theme(axis.title.x = element_text(size = 14,face="bold"),
        axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(hjust = 0.5,size=16,face=20)) + xlab("")

bp_up <- dotplot(filter(ck_bp, NES>1), by="NES") + ggtitle("GO BP Pathways Upregulated") + 
  theme(axis.title.x = element_text(size = 14,face="bold"),
        axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(hjust = 0.5,size=16,face=20)) + xlab("")

bp_down <- dotplot(filter(ck_bp, NES<1), by="NES") + ggtitle("GO BP Pathways Downregulated") + 
  theme(axis.title.x = element_text(size = 14,face="bold"),
        axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(hjust = 0.5,size=16,face=20)) + xlab("")


png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/Pathways_combined_AD_MS_PD_CC.png", res = 300, width = 3400, height = 2200)
cowplot::plot_grid(cc_up,cc_down,align = "v",ncol= 2)
dev.off()

png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/Pathways_combined_AD_MS_PD_MF.png", res = 300, width = 3400, height = 2200)
cowplot::plot_grid(mf_up,mf_down,align = "v",ncol= 2)
dev.off()


png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/Pathways_combined_AD_MS_PD_BP.png", res = 300, width = 3400, height = 2200)
cowplot::plot_grid(bp_up,bp_down,align = "v",ncol= 2)
dev.off()


## Look at genes that drive shared pathways

get_overlapping_genes_per_pathway <- function(ck, pathway, disease1, disease2) {
  
  list1 <- as.data.frame(ck) %>% dplyr::filter(Description == pathway & Cluster == disease1) %>% dplyr::select(core_enrichment)
  list1 <- list1$core_enrichment
  list1 <- strsplit(list1, "/")[[1]]
  
  list2 <- as.data.frame(ck) %>% dplyr::filter(Description == pathway & Cluster == disease2) %>% dplyr::select(core_enrichment)
  list2 <- list2$core_enrichment
  list2 <- strsplit(list2, "/")[[1]]
  
  out <- list()
  out[[disease1]] <- list1
  out[[disease2]] <- list2
  out[["intersect"]] <- intersect(list1, list2)
    return(out)
  
}

muscle_contraction_ad_ms <-  get_overlapping_genes_per_pathway(ck_bp, "muscle contraction", "AD", "MS")

transport_ad_ms <-  get_overlapping_genes_per_pathway(ck_bp, "regulation of transmembrane transport", "AD", "MS")

protein_folding_ms_pd <- get_overlapping_genes_per_pathway(ck_bp, "protein folding", "MS", "PD")
```

Get disease 4-ways

```{r}

plot_disease_4way_and_get_data <- function(res1, res2, disease1, disease2) {

res1 <- as.data.frame(res1) %>% dplyr::mutate(diffexpressed=ifelse(dl_mu>=0.5 & adj_metap_up<=0.05 & n_up>= max(n_tested)/2, "up", 
ifelse(dl_mu<=-0.5 & adj_metap_down<=0.05 & n_down>= max(n_tested)/2, "down", "unchanged"))) 
                                                                      
res2 <- as.data.frame(res2) %>% dplyr::mutate(diffexpressed=ifelse(dl_mu>=0.5 & adj_metap_up<=0.05 & n_up>= max(n_tested)/2, "up", 
ifelse(dl_mu<=-0.5 & adj_metap_down<=0.05 & n_down>= max(n_tested)/2, "down", "unchanged")))                                    

res_merged <- merge(res1,res2,by="symbol")

res_merged <- res_merged %>% dplyr::mutate(sig=
ifelse(diffexpressed.x %in% c("up", "down") & diffexpressed.y %in% c("up", "down"), "Significant in both", 
ifelse(diffexpressed.x %in% c("up", "down") & diffexpressed.y == "unchanged", paste("Changed only in ", disease1, sep=""), 
ifelse(diffexpressed.y %in% c("up", "down") & diffexpressed.x == "unchanged", paste("Changed only in ", disease2, sep=""), "Unchanged in both"))))

res_merged <- res_merged  %>% dplyr::select(dl_mu.x,dl_mu.y,diffexpressed.x, diffexpressed.y,sig,symbol) %>% unique %>% drop_na()
  
  p <- ggplot(res_merged, aes(x=dl_mu.x, y = dl_mu.y, color=sig)) +
  geom_point() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text = element_text(size = 12, colour = "black"), axis.title = element_text(size = 14)) +
  scale_color_manual(values=c("black",  "deepskyblue", "forestgreen", "gray67"), name = "") + geom_hline(yintercept =0) + 
  geom_vline(xintercept=0) + 
  geom_text_repel(data = res_merged[res_merged$sig == "Significant in both",],size=3, aes(x=dl_mu.x,y=dl_mu.y,label=symbol),show.legend = FALSE,box.padding = 0.8, max.overlaps = 20, fontface = "italic") +  xlab(paste("LogFC in ", disease1, sep="")) + ylab(paste("LogFC in ", disease2, sep="")) + guides(fill=guide_legend(title=NULL))
  
  list_out <- list()
  list_out[[1]] <- res_merged
  list_out[[2]] <- p
  return(list_out)
}


ad_pd_4way <- plot_disease_4way_and_get_data(res_ad, res_pd, "AD", "PD")
png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/AD_PD_4_way.png", res = 300, width = 2000, height = 1200)
ad_pd_4way[[2]]
dev.off()
cor(ad_pd_4way[[1]]$dl_mu.x, ad_pd_4way[[1]]$dl_mu.y) ##[1] 0.2820936


ad_ms_4way <- plot_disease_4way_and_get_data(res_ad, res_ms, "AD", "MS")
png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/AD_MS_4_way.png", res = 300, width = 2000, height = 1200)
ad_ms_4way[[2]]
dev.off()
cor(ad_ms_4way[[1]]$dl_mu.x, ad_ms_4way[[1]]$dl_mu.y) ##[1] 0.03540011


pd_ms_4way <- plot_disease_4way_and_get_data(res_pd, res_ms, "PD", "MS")
png("/gne/web/dev/apache/htdocs/people/novikovg/Astrocytes_meta/Harmony_AD_MS_PD_integration_2024/harmony_results_with_removing_mito_clusters/PD_MS_4_way.png", res = 300, width = 2000, height = 1200)
pd_ms_4way[[2]]
dev.off()
cor(pd_ms_4way[[1]]$dl_mu.x, pd_ms_4way[[1]]$dl_mu.y) ##[1] 0.2821504


ad_pd_upregulated <- ad_pd_4way[[1]] %>% dplyr::filter(sig == "Significant in both" & diffexpressed.x == diffexpressed.y) %>% dplyr::filter(dl_mu.x>0)
ad_ms_upregulated <- ad_ms_4way[[1]] %>% dplyr::filter(sig == "Significant in both" & diffexpressed.x == diffexpressed.y) %>% dplyr::filter(dl_mu.x>0)
pd_ms_upregulated <- pd_ms_4way[[1]] %>% dplyr::filter(sig == "Significant in both" & diffexpressed.x == diffexpressed.y) %>% dplyr::filter(dl_mu.x>0)



```


